---
title: "Untitled"
output: html_document
date: "2023-09-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library("readxl")
```

```{r}
setwd("C:/Users/deani/Desktop/Specialeprojekt")
getwd()
```
## Loading the results from the BLAST of ASV_1 against the strain collection

```{r}
setwd("C:/Users/deani/Desktop/Specialeprojekt/BLASTresfiles")
result <- read_tsv("ASV_1_blastres_all.tsv", col_names = F)
head(result)
```

```{r}
# Name columns
df <- result %>% rename(query_ID = X1, target_ID = X2, identity = X3, length = X4, mismatch = X5, gap_openings = X6, query_start = X7, query_end = X8, target_start = X9, target_end = X10, E_value = X11, bit_score = X12)
head(df)
```

```{r}
# Create a column containing strain name and sort columns
df <- df %>% mutate(strain = str_extract(target_ID, pattern = "22.*-[0-9]{1,2}"))
df <- df %>% relocate (E_value) %>% relocate(bit_score) %>% relocate(identity) %>% relocate(strain)
head(df)
```

## Dereplication of hits from same strain

```{r}
dfu <- df %>% distinct(strain, .keep_all = TRUE) # keeps one hit per strain
dfu2 <- df %>% distinct(strain, identity, .keep_all = TRUE) # dereplicates only if hits from same strain has different identities

# count number of rows in the three different dfs
dim(df)[1]
dim(dfu)[1]
dim(dfu2)[1]
```
# how many 100% identity hits in each dataframe

```{r}
sum(dfu$identity == 100)
sum(dfu2$identity == 100)
sum(df$identity == 100)
```
# quick statistics

Original dataframe (not dereplicated)

```{r}
print(paste("Hits with E-val < 10:", dim(df)[1]))
print(paste(">97% identity:", sum(df$identity > 97)))
print(paste("100% identity:", sum(df$identity == 100)))
```
Dereplicated:

```{r}
print(paste("Hits with E-val < 10:", dim(dfu)[1]))
print(paste(">97% identity:", sum(dfu$identity > 97)))
print(paste("100% identity:", sum(dfu$identity == 100)))
```
## Plotting the distribution of BLAST hits

# Histogram of all hits

```{r}
p <- ggplot(df, aes(x=identity)) + geom_histogram(binwidth=0.5, fill = 'grey', color = "white") + scale_x_reverse()  + labs(x ='% Identity', y='Number of hits', title = 'ASV_1 BLAST against strain collection genomes')
p
```
# Histogram of dereplicated hits

```{r}
p1 <- ggplot(dfu, aes(x=identity)) + geom_histogram(binwidth=0.5, fill = 'red', color = "black") + scale_x_reverse() + labs(x ='% Identity', y='Number of hits', title = 'ASV_1 BLAST against strain collection genomes (dereplicated)')
p1
```

## Adding the taxonomic classification by gtdbtk

```{r}
tax <- read_tsv("gtdbtk.bac120.summary.tsv", col_names = TRUE) # load the data
head(tax)
```

# Splitting the classification string into columns

```{r}
t <- c("domain", "phylum", "class", "order", "family", "genus", "species")
taxon <- tax %>% dplyr::select(user_genome, classification) %>% separate(classification, into = t, sep = ";") %>% dplyr::rename("strain" = "user_genome")
taxon
```

# Overview of info in taxonomy table

```{r}
taxon_dropna <- drop_na(taxon)
print(paste("Strains in total:", dim(taxon)[1]))
print(paste("Assigned to some level of taxonomy:", dim(taxon_dropna)[1]))
print(paste("Assigned to genus level:", sum(taxon_dropna$genus != "g__")))
print(paste("Assigned to species level:", sum(taxon_dropna$species != "s__")))
```
# How many are classified as sphingomonads?

```{r}
sphing <- sum(taxon_dropna$genus == "g__Sphingomonas")
not_sphing <- sum(taxon_dropna$genus != "g__Sphingomonas")

print(paste("Number of sphingomonads: ", sphing))
print(paste("Percent sphingomonads: ", sphing/not_sphing*100))
```
# Bar plot of genera in the strain collection

```{r}
#taxon$genus <- sub("g__", "", taxon$genus)

p2 <- ggplot(data=taxon, aes(x=genus)) +
  geom_bar(stat="count", color = "black") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust = .4, size=15),
        axis.title = element_text(size=15)) +
  labs(x ='Genus', y='Number of isolates')

p2

```

# Add taxonomy info to BLAST hit df

```{r}
ndf <- left_join(dfu, taxon, by="strain")
head(ndf)
```

## Colour histogram of hits by taxonomy

```{r}
ndf$genus <- sub("g__", "", ndf$genus)
```


```{r}
# histogram of identity coloured by genus
p3 <- ggplot(ndf, aes(x=identity, fill=genus)) +
  theme_bw() +
  geom_histogram(binwidth=1, colour="grey30") +
  labs(x ='Percent identity', y='Number of isolates', fill="Genus") +
  guides(fill=guide_legend(ncol=2)) +
  theme(axis.title = element_text(size=12))
p3
```


```{r}
# bar plot of mismatches coloured by genus
p3.2 <- ggplot(data=ndf, aes(x=mismatch, fill=genus)) +
  geom_bar(stat="count")
p3.2
```


```{r}
# histogram of identity coloured by species
p4 <- ggplot(ndf, aes(x=identity, fill=species)) + geom_histogram(binwidth=0.5, color = "black") + scale_x_reverse() + labs(x ='% Identity', y='Number of hits', title = 'ASV_1 BLAST coloured by species') + xlim(101,90) # add x axis limit
p4

# bar plot of mismatches coloured by species
p4.2 <- ggplot(data=ndf, aes(x=mismatch, fill=species)) + geom_bar(stat="count") + xlim(-1,20) # add x axis limit
p4.2
```

# Produce a list of sphingomonas genomes file names

```{r}
sphingos <- filter(taxon, genus == "g__Sphingomonas") %>% mutate(file = paste0(strain, ".fasta")) %>% dplyr::select(file)
head(sphingos)
write.table(sphingos, file = "sphingomonads.txt", sep = "\t", row.names = FALSE, col.names = FALSE, quote = FALSE)
```
# Produce a list of genomes with 100% identity to ASV_1 file names

```{r}
ndf100 <- ndf %>% filter(identity == 100) %>% mutate(file = paste0(strain, ".fasta"))
ndf100_list <- ndf100  %>% dplyr::select(file)
dim(ndf100_list)
write.table(ndf100_list, file = "no_mismatch_hits.txt", sep = "\t", row.names = FALSE, col.names = FALSE, quote = FALSE)
```

```{r}
ndf1001 <- ndf100
ndf1001$species <- sub("s__", "", ndf1001$species)
ndf1001 <- ndf1001 %>% separate(species, into=c("gen", "species"), sep=" ")
#ndf100$species <- sub("", "Unknown species", ndf100$species)
```

```{r}

```

```{r}
ggplot(ndf100 %>% filter(!is.na(species))) +
  geom_bar(aes(x=species)) +
  theme_bw() +
  labs(x ='Species', y='Number of isolates') +
  theme(axis.title = element_text(size=14),
        axis.text.x = element_text(angle = -90, hjust=0, vjust = .3, size=12))

```


## Adding sequencing and assembly quality info

#Reading table and assigning meaningful column names
```{r}
info <- read_excel("Overview_of_all_isolates_2022.xlsx")
info <- info %>% rename(genome_size=`Genome size (Mb)`, strain=Strain, N50_reads=`N50 reads`, coverage=`Avg cov`, ONT_bc=`ONT bc`, DNA_conc=`DNA conc. used in lib (ng/µl)`, read_counts=`Read counts`, filtered_data=`Filtered_data (Mb)`, median_reads = `Median reads (bp)`, mean_reads = `Mean reads (bp)`, read_len_stdev = `Read Len. stdev`, N50_assembly=`N50 assembly (bp)`, lineage=`Lineage (tentative)`) %>% filter(!is.na(genome_size))
#head(info)
```

# Adding information to df
```{r}
ndf100 <- left_join(ndf100,info, by="strain")
#ndf100
```

# Adjusting the genome size to bp from Mb
```{r}
ndf100 <- ndf100 %>% mutate(genomesize = case_when(genome_size < 15 ~ genome_size*1000000, genome_size > 15 ~ genome_size))
#ndf100$genome_size
#ndf100$genomesize
```

# Descriptive statistics of assembly quality markers
```{r}
summary(ndf100$genomesize)
quantile(ndf100$genomesize, 0.1)
quantile(ndf100$genomesize, 0.9)
summary(ndf100$Contigs)
summary(ndf100$coverage)
summary(ndf100$N50_assembly)
```
With plots 

```{r}
pl1 <- ggplot(ndf100, aes(x=genomesize))+geom_histogram(binwidth=50000)
pl2 <- ggplot(ndf100, aes(x=Contigs))+geom_histogram(binwidth=2)
pl3 <- ggplot(ndf100, aes(x=coverage))+geom_histogram(binwidth = 5)
pl4 <- ggplot(ndf100, aes(x=N50_assembly))+geom_histogram(binwidth=50000)

pl1
pl2
pl3
pl4
```



## Filtering the genomes with 100% identity using assembly quality

```{r}
filtered <- ndf100 %>% filter(Contigs <= 7) %>% filter(coverage >= 13) %>% filter(N50_assembly >= 3470091) %>% filter(genomesize >= 4060099) %>% filter(genomesize <= 7543137) 

dim(filtered)

#filtered <- ndf100 %>% filter(Contigs <= 10) %>% filter(coverage > 20) %>% filter(N50_assembly > 50000) %>% arrange(desc(coverage))
#filtered

```
# Writing the resulting list of genomes to a file

```{r}
#filtered_list <- filtered %>% mutate(file = paste0(strain, ".fasta")) %>% select(file)
#write.table(filtered_list, file = "filtered.txt", sep = "\t", row.names = FALSE, col.names = FALSE, quote = FALSE)

filtered_list <- filtered2 %>% mutate(file = paste0(strain, ".fasta")) %>% select(file)
filtered_list
write.table(filtered_list, file = "filtered.txt", sep = "\t", row.names = FALSE, col.names = FALSE, quote = FALSE)
```

## Adding information of lengths of individual contigs (the .fasta.fai files combined)

```{r}
index <- read.table("index.tsv", header = FALSE) # read the file

n <- c("contig", "contig_nr", "strain") # columns to separate contig info into
index <- index %>% separate(V1, into = n, sep = "_") %>% rename(contig_length = V2) %>% select(contig_nr, strain, contig_length)
#head(index)
```

```{r}
# sort after contig length and keep one contig per strain: creates a list of the longest contig from each assembly
index <- index %>% arrange(desc(contig_length)) %>% distinct(strain, .keep_all = TRUE) %>% rename(longest_contig = contig_length)
#head(index)
```


```{r}
ndf100 <- left_join(ndf100,index, by="strain")
```

## Filtering for minimum length of the largest contig
The largest contig should be genome-sized (doesn't remove any rows after the previous filtering)

```{r}
dim(ndf100 %>% filter(Contigs <= 10) %>% filter(coverage > 20) %>% filter(N50_assembly > 50000) %>% filter(longest_contig>3000000))
```



